Transcriptomic Analysis of Expressions of Ten genes and Their Prognostic Value in Head and Neck Squamous Cell Carcinoma

### 2021-05-09 Cover_letter_JPM.tex
modified Figure 1
Updates from Overleaf

April 30, 2021

@@

%{%
1) % {2-4}
"Please shortly explain the TCGA database in the Introduction and summarize the advantages of its use."\\
We re-wrote the Introduction in detail according to the suggestion by reviewer \#2. %(comment No 2-4).
%}

2) % {3-1}
"Are the conclusions supported by the results?
Only computer computing in this research to support 20 candidate biomarkers, ... are all heavily associated with the prognosis of OS.
There is little information provided. The authors should provide more evidence to support this article."\\
We provide more evidence by additional analysis of TCGA and GEO databases. Moreover, we reviewed literature that supports the prognostic impact of these candidate genes.
According to the suggestion by reviewer \#3,
we re-wrote the Discussion with additional information, and
%}
%{% Answer_3-1_USP10.pdf, Answer_2-1.pdf
%According to the reviewers' comments, we 
added new supplementary Figure S1, S2, and S3 %, S2, S3, and Table S2. We also replaced Figure 3D and 4A with high magnitude images from the same slides. We transferred the original Figure 2 into supplementary Figure S1 
to provide more evidence from the significant literature review.
%}

@@

We appreciate the reviewer for kind instruction of expression cutoff in detail.\\
We modified the workflow as ...

[after cut-off finding]
Adjustment of multiple test p-values should be applied here, which is missing in the current manuscript.

***=> Bonferroni correction should be used in final decision.
%位置要換一下：
The false discovery rate
645 (FDR) (< 0.05) could be used to pick up the optimal P value to ensure the control for
646 type I error of the Kaplan-Meier survival test during the cutoff finding procedure.
4.2. Cutoff Finder Core Engine.

@@

We appreciate the reviewer for the critical comment.
We change new Title, abstract, change tone

@@

We appreciate the reviewer for the critical comments and thanks for the agreement of our workflow in cancer research.
調整 workfow 更加嚴格
以增加 validation rate
The p-values are affected by the sample size, which has be described in the main text.
as following:
% ???
% x
%how to apply or modify the R scripts to use on other datasets
%\begin{itemize} %summarize the elements of the workflow
\begin{outline}
The data from TCGA database or other databases could be applied by this workflow, since modification of the R script is necessary to adapt the difference structure among datasets (e.x. data schema, table size, definition of features, number of features and their location on the table column).\\
In short, the elements of the workflow with the \{codes\} shows:\\
(please also see Figure \ref{fig:workflow})
\1 $[\# step 1 (main)]$
\{main\_marginSFP\_HNSCC.R\}
        \2 dataset retrieval from TCGA GDC data portal (in GDAC format)
%\1 $[\# step 1]$
    \2 \{TCGA\_HNSCC\_marginSFP.R\}
        \3 \underline{raw data cleaning}
        \3 \underline{data transformation and merge}
        \3 \underline{dichotomization of clinical features}
%     \item Using a serial cut from 30\% to 70\% percentile of the cohort, it could find the least P-value cutoff of the Kaplan-Meier analysis for each gene expression.
        \3 survival analysis
            \4 cut off finder by \{cutofFinder\_func\_HNSCC.R\}
%     \item The R script program could automatically scan protein-coding genes and generate the Kaplan-Meier plots and Cox's univariate/multivariate tables.
            \4 Cox modelling
\1 $[\# step 2]$
\{analysis\_export.R\}
    \2 analysis of the tables generated from \# step 1
    \2 candidate selection
%    \item Our analysis could discover the pronounced biomarkers, which impact HNSCC's survival under the stringent Bonferroni adjustment.
\1 (remark: only the \underline{underlined} portion in \{TCGA\_HNSCC\_marginSFP.R\} should be re-coded aggressively)
\end{outline}

@@

We appreciate the reviewer for the critical comment and modified the Introduction accordingly.
= 詳細描述 univariable and multivariable tests work
In a multivariable test, which variables were used? 
= 列出來，並且強調 surgical margin

@@

In Figure 2, unlabeled scales on the x-axis???
除去 unlabeled stick 即可立刻改進
\begin{MyColorPar}{red}

@@

\begin{outline} % still blue
In conclusion, 20 candidate biomarkers, DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, DKC1, as well as ZNF557, ZNF266, IL19, MYO1H, FCGBP, LOC148709, EVPLL, PNMA5, IQCN (former name as KIAA1683), and NPB, are associated with the prognosis of OS in our study.
Comparison with previous publications or database analyses,

>>>>

May 16, 2021

@@

%{%
Reviewer 1: % {2-4}
"Please shortly explain the TCGA database in the Introduction and summarize the advantages of its use."\\
We re-wrote the Introduction in detail according to the suggestion by reviewer \#2. %(comment No 2-4).
%}

Reviewer 2: % {3-1}
"Are the conclusions supported by the results?
Only computer computing in this research to support 20 candidate biomarkers, ... are all heavily associated with the prognosis of OS.
There is little information provided. The authors should provide more evidence to support this article."\\
We provide more evidence by additional analysis of TCGA and GEO databases. Moreover, we reviewed literature that supports the prognostic impact of these candidate genes.
According to the suggestion by reviewer \#3,
we re-wrote the Discussion with additional information, and
%}
%{% Answer_3-1_USP10.pdf, Answer_2-1.pdf
%According to the reviewers' comments, we 
added new supplementary Figure S1, S2, and S3 %, S2, S3, and Table S2. We also replaced Figure 3D and 4A with high magnitude images from the same slides. We transferred the original Figure 2 into supplementary Figure S1 
to provide more evidence from the significant literature review.
%}

@@

We appreciate the reviewer for kind instruction of expression cutoff in detail.\\
*We modified the workflow as ...

We modified the last paragraph of "Cutoff Finder Core Engine" in the Material and Methods section.
The paragraphs marked in \textcolor{red}{Red} from page 2 (line 77) to page 3 (line 129) are the part of the manuscript that has been revised. Also, it was shown here in the following paragraph.
\\[0.3cm]
%as follows:
%\begin{MyColorPar}
The \acrfull{fdr} ($< 0.05$) correction has been applied to ensure the control of type I error of multiple test \textit{P} values during our cutoff finding procedure.
%\end{MyColorPar}

%***=> Bonferroni correction should be used in final decision.

@@

We appreciate the reviewer for the critical comment.
We changed title as "Transcriptomic Analysis of Expressions of Ten genes and Their Prognostic Value in Head and Neck Squamous Cell Carcinoma", and abstract; we also changed the tone in our manuscript to focus on the identified biomarkers in the head and neck squamous cell carcinoma.

* more evidence to support or validate them

@@

We appreciate the reviewer for the critical comments and thanks for the agreement of our workflow in cancer research.
***調整 workfow 更加嚴格以增加 validation rate
*The p-values are affected by the sample size, which has be described in the main text.
as following:

*the authors should look for more data to verify these biomarkers.
% ???

% x
\begin{outline}
The data from TCGA database or other databases could be applied by this workflow, since modification of the R script is necessary to adapt the difference structure among datasets (e.x. data schema, table size, definition of features, number of features and their location on the table column).\\
In short, the elements of the workflow with the \{codes\} shows:\\
(please also see Figure \ref{fig:workflow})
\1 $[\# step 1 (main)]$
\{main\_marginSFP\_HNSCC.R\}
        \2 dataset retrieval from TCGA GDC data portal (in GDAC format)
%\1 $[\# step 1]$
    \2 \{TCGA\_HNSCC\_marginSFP.R\}
        \3 \underline{raw data cleaning}
        \3 \underline{data transformation and merge}
        \3 \underline{dichotomization of clinical features}
%     \item Using a serial cut from 30\% to 70\% percentile of the cohort, it could find the least P-value cutoff of the Kaplan-Meier analysis for each gene expression.
        \3 survival analysis
            \4 cut off finder by \{cutofFinder\_func\_HNSCC.R\}
%     \item The R script program could automatically scan protein-coding genes and generate the Kaplan-Meier plots and Cox's univariate/multivariate tables.
            \4 Cox modelling
\1 $[\# step 2]$
\{analysis\_export.R\}
    \2 analysis of the tables generated from \# step 1
    \2 candidate selection
%    \item Our analysis could discover the pronounced biomarkers, which impact HNSCC's survival under the stringent Bonferroni adjustment.
\1 (remark: only the \underline{underlined} portion in \{TCGA\_HNSCC\_marginSFP.R\} should be re-coded aggressively)
\end{outline}

@@

We appreciate the reviewer for the critical comment and modified the Introduction accordingly.
= 詳細描述 univariable and multivariable tests work in the Cox proportional hazard regression model.
In a multivariable test, which variables were used? 
= 列出來，並且強調 surgical margin

@@

In Figure 2, the ticks and labels on the x-axis were corrected.

@@

\begin{outline} % highlight still blue
In conclusion, 20 candidate biomarkers, DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, DKC1, as well as ZNF557, ZNF266, IL19, MYO1H, FCGBP, LOC148709, EVPLL, PNMA5, IQCN (former name as KIAA1683), and NPB, are associated with the prognosis of OS in our study.
Comparison with previous publications or database analyses,

### 2021-05-09 main.tex
Updates from Overleaf

%\label{key:cutoff}
To evaluate the effect of gene expression on the patient's survival, we introduced the stratifying of patients with Kaplan-Meier survival analysis according to each gene's low/high expression.
Our cutofFinder\_func subroutine employs the minimum \textit{P} value approach to recognizing cutoff points in continuous gene expression measurement for patients sub-population.
First, patients were ordered by \acrshort{rnaseq} value (RSEM) of a given gene. Next, patients were stratified at a serial cut (counted by person ranked in 30\% to 70\% percentile of the cohort; please see Figure \ref{fig:figure1} cutoff engine). The survival risk differences of the two groups were estimated by log-rank test to yield around 165 Kaplan-Meier \textit{P} values for each gene.
Then, the optimal cutoff of \acrshort{rnaseq}, giving the minimum \textit{P} value, was selected by the cutofFinder\_func subroutine.
This iteration method could calculate all possible cutoff of each gene expression in this cohort. At each run of cutofFinder\_func function call for an individual gene, it returned an optimal cutoff value (e.x. 0.027 for gene \acrlong{CAMK2N1}, \acrshort{CAMK2N1}, Figure \ref{fig:figure4}). The optimal cutoff value and its correlated patient grouping size (e.x. low-expression in 262 persons vs. high-expression in 152 persons with gene \acrshort{CAMK2N1}) would be returned to the main program to allow downstream Cox survival analysis. The percentile range we applied as 30\% to 70\% was used to avoid a small grouping effect\cite{Miller1982}\cite{Mizuno2009a}.
In case there was no significant \textit{P} value, a median expression of this gene was set as its cutpoint as usual.

\subsection{Statistical Consideration}

Our workflow has loops to call function survival\_marginSFP(GeneName) with given GeneName to process the survival analysis gene by gene.
We dichotomized the clinicopathological features, which includes gender (male/ female), age at diagnosed (below/beyond 65 year-old), clinical tumor size (T1-2/T3-4), clinical nodal status (negative/positive), clinical distant metastasis (negative/positive), TNM staging (early/late), surgical margin status (negative/positive) and tobacco exposure (low/high). The patients were grouped by an \acrshort{rnaseq} value of each gene, cut at low- or high-expression on an optimal \textit{P} value obtained from the cutofFinder\_func subroutine (see the section of "Cutoff Finder Core Engine"). Pearson's chi-square test was used for these binary variables. Kaplan-Meier survival was analyzed using the log-rank test for two groups OS comparison.
The Cox proportional hazards regression is the widely accepted approach for modeling survival while accounting for confounding factors\cite{Magen2019}. Univariate and multivariate Cox proportional regression model\cite{Andersen1982}, using the "coxph" function in R package "survival", was applied to calculate hazard ratio, \acrfull{ci95} and its significance, and to estimate the independent contributions of each clinicopathological features to the OS.
Results were considered statistically significant when a two-sided \textit{P} $<$ 0.05, or a lower threshold if indicated.
The \acrfull{fdr} ($< 0.05$) could be used to pick up the optimal \textit{P} value to ensure the control for type I error of the Kaplan-Meier survival test during the cutoff finding procedure.
%statistical inference problems in biomedical and genomic data analysis routinely involve the simultaneous test of thousands, or even millions, of null hypotheses. Examples
There were also multiple correlated tests of null hypotheses in the Kaplan-Meier family during our global scanning of protein-coding genes. The stringent Bonferroni correction could result in an adjusted \textit{P} value to ensure the control for type I error. 
%There were multiple correlated tests in the family of Kaplan-Meier survival hypotheses during our global scanning of protein-coding genes. The stringent Bonferroni correction could result in an adjusted P-value to ensure the control for type I error.

>>>>

%\label{key:cutoff}
To evaluate the effect of gene expression on the patient's survival, we introduce a sliding window cutoff selection by stratifying patients with Kaplan-Meier survival analysis according to each gene's low/high expression.
Our cutofFinder\_func subroutine employs the minimum \textit{P} value approach to recognizing cutoff points in continuous gene expression measurement for patients sub-population.
First, patients were ordered by \acrshort{rnaseq} value (RSEM) of a given gene. Next, patients were stratified at a serial cut (counted by person ranked in 30\% to 70\% percentile of the cohort; please see Figure \ref{fig:figure1} cutoff engine). The survival risk differences of the two groups were estimated by log-rank test to yield around 165 Kaplan-Meier \textit{P} values for each gene.
Then, the optimal cutoff of \acrshort{rnaseq}, giving the minimum \textit{P} value, was selected by the cutofFinder\_func subroutine.
This iteration method could calculate all possible cutoff of each gene expression in this cohort. At each run of cutofFinder\_func function call for an individual gene, it returned an optimal cutoff at specific patient grouping  (e.x. low-expression in 262 persons vs. high-expression in 152 persons with gene \acrlong{CAMK2N1}, \acrshort{CAMK2N1}, please see Figure \ref{fig:figure4}). The cutoff would be returned to the main program to allow downstream Cox survival analysis. The percentile range we applied as 30\% to 70\% was used to avoid a small grouping effect\cite{Miller1982}\cite{Mizuno2009a}.
In case there was no significant \textit{P} value, a median expression of this gene was set as its cutpoint as usual.
\begin{MyColorPar}{red}
The \acrfull{fdr} ($< 0.05$) correction has been applied to ensure the control of type I error of multiple test \textit{P} values during our cutoff finding procedure.
\end{MyColorPar}

\subsection{Statistical Consideration for Survival Analysis}

Our workflow has loops to call function survival\_marginSFP(GeneName) with given GeneName to process the survival analysis gene by gene.
We dichotomized the clinicopathological features, which includes gender (male/ female), age at diagnosed (below/beyond 65 year-old), clinical tumor size (T1-2/T3-4), clinical nodal status (negative/positive), clinical distant metastasis (negative/positive), TNM staging (early/late), surgical margin status (negative/positive) and tobacco exposure (low/high). The patients were grouped by an \acrshort{rnaseq} value of each gene, cut at low- or high-expression on an optimal \textit{P} value obtained from the cutofFinder\_func subroutine (see the section of "Cutoff Finder Core Engine"). Pearson's chi-square test was used for these binary variables. Kaplan-Meier survival was analyzed using the log-rank test for two groups OS comparison.

% reviewer comment 
\begin{MyColorPar}{red}
%The Cox proportional hazards regression is the widely accepted approach for modeling survival while

The Cox proportional-hazards model\cite{Cox1972} is commonly used for modeling survival analysis data. It allows to analyze survival with respect to one or more predictor variables and to provides the effect size (hazard) for them\cite{Bradburn2003b}. % Classification trees and artificial neural networks in survival analysis
The Cox model is also accounting for confounding factors\cite{Magen2019}.
It is expressed by the hazard function denoted by h(t). Briefly, the hazard function can be interpreted as the risk of a specific event (e.x. death) at time t. It can be estimated as follow:

%Linear regression is useful as a concept in biomarker discovery.
%A linear regression equation that relates a Y-variable to n X-variables is written as\\[1cm]
%\par
%Highlights
% https://quicklatex.com
\begin{flushleft}
% equation
%\begin{math}

%ground truth Y (e.x. patient's survival):\\[0.5cm]
$h(t) = h_0(t) \times exp(\beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + ... + \beta_n X_n)$\\[0.3cm]
% natural exponential function => y=e^x
\end{flushleft}
%Y$:
%$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + ... + \beta_n X_n + \epsilon$
%\linebreak
%\linebreak
where,\\
\begin{outline} % for highlight
\1 t represents the survival time
\1 $h(t)$ is the hazard function determined by a set of n covariates ($X_1...X_n$)
\1 the coefficients ($\beta_1...\beta_n$) measure the impact (i.e., the effect size) of covariates
\1 the term $h_0$ is called the baseline hazard. It corresponds to the value of the hazard if all the $X_i$ are equal to zero (the quantity exp(0) equals 1). The ‘t’ in $h(t)$ indicates us that the hazard may vary over time

\end{outline}

Input $X_1...X_n$ (e.x. patients features: such as age, gender, gene expression, cancer stage, or even spiritual, emotional, and social status)
\\[0.5cm]
Coefficients $\beta_0...\beta_n$ (built in model)\\[0.5cm]
%A residual (error) term is calculated as $\epsilon=Y - \hat Y$\\
%Since the model established by survival analysis, predicted value $\hat Y$ is calculated as:\\[0.3cm]
%$\hat Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + ... + \beta_n X_n$
%\linebreak
%(predicted Y, output) intercept, slope, independent (input) \Times
%\end{math}

Thus, the biomarker discovery strategy is survival modeling through collections of $X_1...X_n$ features from cancer datasets.\\[1cm]

Univariate and multivariate Cox proportional regression model\cite{Andersen1982}, using the "coxph" function in R package "survival", was applied to calculate hazard ratio, \acrfull{ci95} and its significance, and to estimate the independent contributions of each clinicopathological features to the OS.

In a multivariate test, those variables used include the clinicopathological features (gender, age at diagnosed separated by 65 year-old, clinical tumor size [T1-2/T3-4], clinical nodal status [N0/N+], clinical distant metastasis [M0/M1], TNM staging [stage 1-2/stage 3-4], surgical margin status [negative/positive] and tobacco exposure [low/high]), and gene expression level [low/high] defined by the cutoff.
\end{MyColorPar}
%%

Results were considered statistically significant when a two-sided \textit{P} $<$ 0.05, or a lower threshold if indicated.
The \acrfull{fdr} ($< 0.05$) could be used to pick up the optimal \textit{P} value to ensure the control for type I error of the Kaplan-Meier survival test during the cutoff finding procedure.
%statistical inference problems in biomedical and genomic data analysis routinely involve the simultaneous test of thousands, or even millions, of null hypotheses. Examples
There were also multiple correlated tests of null hypotheses in the Kaplan-Meier family during our global scanning of protein-coding genes. The stringent Bonferroni correction could result in an adjusted \textit{P} value to ensure the control for type I error. 
%There were multiple correlated tests in the family of Kaplan-Meier survival hypotheses during our global scanning of protein-coding genes. The stringent Bonferroni correction could result in an adjusted P-value to ensure the control for type I error.
